Syntax and Semantics of Abstract Binding Trees

### 2015-11-27 paper.tex
adjust notation for renaming operators

Declaring a new assignable consists in providing an initial value, and an
expression binding a symbol (which shall represent the assignable in scope).
Note that the functoriality of $\Operators$ guarantees for any renaming
$\Of{\varrho}{\Upsilon\inj\Upsilon'}$ and $\Of{f}{a\ra a'}$, a lifted map
$\Of{\FLift{\varrho}{f}}{\Operators\Pair{\Upsilon}{a}\inj\Operators\Pair{\Upsilon'}{a'}}$
such that $\IsOperator{\Upsilon'}{\FLift{\varrho}{f}(\vartheta)}{a'}$ when
$\IsOperator{\Upsilon}{\vartheta}{a}$. Because the category
$\Discrete{\Arities}$ is discrete and has only identity arrows, through an abuse
of notation we will often write $\FLift{\varrho}{}$ to mean
$\FLift{\varrho}{\ArrId{}}$. In particular, the renaming
$\Upsilon,u\mapsto\Upsilon,v$ shall take $\mathtt{get}[u]$ to $\mathtt{get}[v]$.

@@

\begin{gather*}
  \infer{
    \IsRenaming{\varrho}{x}{x}
  }{
  }
\\[6pt]
  \infer{
    \IsRenaming{\varrho}{
      \MApp{\MV{m}}{\vec{u}}{M_0,\dots,M_n}
    }{
      \MApp{\MV{m}}{\vec{v}}{N_0,\dots,N_n}
    }
  }{
    \Match{\varrho(\vec{u})}{\vec{v}}
&
    \HypJ{
      \IsRenaming{\varrho}{M_i}{N_i}
    }{
      \Leq{i}{n}
    }
  }
\\[6pt]
  \infer{
    \IsRenaming{\varrho}{
      \App{\vartheta}{E_0,\dots,E_n}
    }{
      \App{\vartheta'}{F_0,\dots,F_n}
    }
  }{
    \Match{\FLift{\varrho}{}(\vartheta)}{\vartheta'}
&
    \HypJ{
      \IsRenaming{\varrho}{E_i}{F_i}
    }{
      \Leq{i}{n}
    }
  }
\\[6pt]
  \infer{
    \IsRenaming{\varrho}{
      \MkBTm{\vec{u}}{\vec{x}}{M}
    }{
      \MkBTm{\vec{u}}{\vec{x}}{N}
    }
  }{
    \IsRenaming{\varrho\setminus\vec{u}}{M}{N}
  }
\end{gather*}

@@

Interpretation of operator applications is the most complicated. Recall that,
unlike in standard treatments of universal algebra, our operators are indexed by
symbol collections; therefore, operators must pass through suitable renamings in
order to be used in the interpretation. Let us begin by constructing for each
operator $\IsOperator{\Upsilon}{\vartheta}{a}$ the morphism
$\SemBrackets{\vartheta}_P$ which shall rename the parameters of the operator
using the environment:
\[
  \begin{tikzcd}[cramped, ampersand replacement = \&]
    |[alias = Src]|
    {\IMode{
      {\SemBrackets{
        \Theta\triangleright\Upsilon\parallel\Gamma
      }}_{P}
    }}
      \arrow[
        , r
        , swap
        , "\IMode{\pi_2}"
      ]
\&
    {\IMode{\SymPsh^{\{\Upsilon\}}}}
      \arrow[
        , r
        , swap
        , "\IMode{\FLift{\pi_{(-)}}{}}"
      ]
\&[1.5em]
    |[alias = Tgt]|
    {\IMode{\Operators_a}}
      \arrow[
        , from = Src
        , to = Tgt
        , bend left
        , dashed
        , "\OMode{\SemBrackets{\vartheta}_{P}}"
      ]
  \end{tikzcd}
\]

>>>>

Declaring a new assignable consists in providing an initial value, and an
expression binding a symbol (which shall represent the assignable in scope).
Note that the functoriality of $\Operators$ guarantees for any renaming
$\Of{\varrho}{\Upsilon\inj\Upsilon'}$ and $\Of{f}{a\ra a'}$, a lifted map
$\Of{\FLift{\varrho}{f}}{\Operators\Pair{\Upsilon}{a}\inj\Operators\Pair{\Upsilon'}{a'}}$
such that $\IsOperator{\Upsilon'}{\FLift{\varrho}{f}(\vartheta)}{a'}$ when
$\IsOperator{\Upsilon}{\vartheta}{a}$. In particular, the renaming
$\Upsilon,u\mapsto\Upsilon,v$ shall take $\mathtt{get}[u]$ to
$\mathtt{get}[v]$. Because the category $\Discrete{\Arities}$ is discrete and
has only identity arrows, through an abuse of notation we will often write
$\vartheta\varrho$ to mean $\FLift{\varrho}{\ArrId{}}(\vartheta)$.

@@

\begin{gather*}
  \infer{
    \IsRenaming{\varrho}{x}{x}
  }{
  }
\\[6pt]
  \infer{
    \IsRenaming{\varrho}{
      \MApp{\MV{m}}{\vec{u}}{M_0,\dots,M_n}
    }{
      \MApp{\MV{m}}{\vec{v}}{N_0,\dots,N_n}
    }
  }{
    \Match{\varrho(\vec{u})}{\vec{v}}
&
    \HypJ{
      \IsRenaming{\varrho}{M_i}{N_i}
    }{
      \Leq{i}{n}
    }
  }
\\[6pt]
  \infer{
    \IsRenaming{\varrho}{
      \App{\vartheta}{E_0,\dots,E_n}
    }{
      \App{\vartheta\varrho}{F_0,\dots,F_n}
    }
  }{
    \HypJ{
      \IsRenaming{\varrho}{E_i}{F_i}
    }{
      \Leq{i}{n}
    }
  }
\\[6pt]
  \infer{
    \IsRenaming{\varrho}{
      \MkBTm{\vec{u}}{\vec{x}}{M}
    }{
      \MkBTm{\vec{u}}{\vec{x}}{N}
    }
  }{
    \IsRenaming{\varrho\setminus\vec{u}}{M}{N}
  }
\end{gather*}

@@

Interpretation of operator applications is the most complicated. Recall that,
unlike in standard treatments of universal algebra, our operators are indexed by
symbol collections; therefore, operators must pass through suitable renamings in
order to be used in the interpretation. Let us begin by constructing for each
operator $\IsOperator{\Upsilon}{\vartheta}{a}$ the morphism
$\SemBrackets{\vartheta}_P$ which shall rename the parameters of the operator
using the environment:
\[
  \begin{tikzcd}[cramped, ampersand replacement = \&]
    |[alias = Src]|
    {\IMode{
      {\SemBrackets{
        \Theta\triangleright\Upsilon\parallel\Gamma
      }}_{P}
    }}
      \arrow[
        , r
        , swap
        , "\IMode{\pi_2}"
      ]
\&
    {\IMode{\SymPsh^{\{\Upsilon\}}}}
      \arrow[
        , r
        , swap
        , "\IMode{\FLift{\pi_{(-)}}{\ArrId{}}}"
      ]
\&[1.5em]
    |[alias = Tgt]|
    {\IMode{\Operators_a}}
      \arrow[
        , from = Src
        , to = Tgt
        , bend left
        , dashed
        , "\OMode{\SemBrackets{\vartheta}_{P}}"
      ]
  \end{tikzcd}
\]
